---
title: "Homework 2"
author: Nancy Fang (nf2466)
output: github_document
---

```{r setup}
library(tidyverse)
library(readxl)
library(readr)
```


## Problem 1
Read the Mr. Trash Wheel data set
```{r}
trashwheel_df=
  read_xlsx("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
  sheet = "Mr. Trash Wheel",
  range = cell_cols("A:N")) %>%
  janitor::clean_names()%>%
  drop_na(dumpster) %>%
    mutate(
      sports_balls = round(sports_balls),
      sports_balls = as.integer(sports_balls)
    )
```

Read in precipitation data for 2017 and 2018

```{r}
precip_2018 = 
  read_xlsx(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet="2018 Precipitation",
    skip = 1,
  )%>%
  janitor:: clean_names() %>%
  drop_na(month)%>%
  mutate(year=2018)%>%
  relocate(year)

precip_2017 = 
  read_xlsx(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet="2017 Precipitation",
    skip = 1,
  )%>%
  janitor:: clean_names() %>%
  drop_na(month)%>%
  mutate(year=2017)%>%
  relocate(year)

```
Now combine annual precipitation

```{r}
precip_df = 
  bind_rows(precip_2018, precip_2017)

```
Switch month variable from number to character
```{r}
month_df = 
  tibble(
    month = 1:12,
    month_name = month.name
  )

left_join(precip_df,month_df, by = "month")

```
This dataset contains information from the Mr. Trash Wheel trash collector in Baltimore, MD. As trash enters the inner harbor, the trash wheel collects trash and stores it in a dumpster. The dataset contains information on year, month, and trash collected, including some specific kinds of trash. There are a total of 
`r nrow(trashwheel_df)` in our final dataset. Additional data sheets include month precipitation data. 

* The median number of sports balls found in a dumpster in 2017 was `r trashwheel_df %>% filter(year == 2017) %>% pull(sports_balls) %>% median()`
* The total precipitation in 2018 was `r precip_df %>% filter(year == 2018) %>% pull(total) %>% sum()` inches.


## Problem 2

Read in Subway Entrance and Exit Data
```{r}
subway_df=
  read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>%
  janitor::clean_names()%>%
  select(line:entry,vending,ada)%>%
  mutate(entry = recode(entry, YES = "TRUE", NO = "FALSE"))%>%
  mutate(entry = as.logical(entry))
  
```
This dataset contains information about subway stations in NYC. It contains information on subway line, station, latitude/longitude, which routes it serves, entrance type, whether there are vending machines, and its ADA compliance. To read this dataset, I imported a csv file, cleaned the names, selected the columns I wanted to keep, and changed the entry variable from a character variable to a logical variable. There are a total of `r nrow(subway_df)` rows and `r ncol(subway_df)` columns. The dataset appears to overall be tidy except for the route variables .

There are `r distinct(subway_df,line,station_name) %>% count()` distinct stations in this dataset. There are `r filter(subway_df, ada == TRUE) %>% distinct(line, station_name) %>% count()` distinct stations that are ADA compliant. The proportion of station entrances/exits with no vending that allow entry is `r filter(subway_df, vending == 'NO' & entry==TRUE) %>%  count() / filter(subway_df, vending == 'NO') %>% count()`.

```{r}
subway_tidy =
  subway_df %>%
  mutate_at(vars(route1:route11), as.character) %>%
  pivot_longer(route1:route11, names_to = "route_name", names_prefix = "route", values_to = "route_number") %>%
  drop_na(route_number)

```
There are `r filter(subway_tidy,route_number=='A')%>%distinct(line,station_name)%>%count()` distinct stations that serve the A train. There are `r filter(subway_tidy,route_number=='A') %>% filter(ada==TRUE)%>%distinct(line,station_name)%>%count()` stations on the A line that are ADA compliant.


Problem 3
---
Read pols-month data
```{r pols csv}
pols_df=
  read_csv("./data/fivethirtyeight_datasets/pols-month.csv")%>%
  janitor::clean_names() %>%
  separate(mon, into = c('year','month','day')) %>%
  mutate(month = month.abb[as.factor(month)]) %>%
  mutate_at(vars(year), as.numeric)%>%
  mutate(president = case_when (
    prez_gop == 1 ~ "gop",
    prez_dem == 1 ~ "dem")) %>%
  subset(select = -c(prez_gop, prez_dem, day)) %>%
  relocate(president, .after=month)
```
Read snp data
```{r snp csv}
snp_df=
  read_csv("./data/fivethirtyeight_datasets/snp.csv")%>%
  janitor::clean_names() %>%
  separate(date, into = c('month','day','year')) %>%
  mutate(month = month.abb[as.factor(month)]) %>%
  mutate_at(vars(year), as.numeric)%>%
  subset(select = -c(day)) %>%
  relocate(year, .before=month)
```
Read unemployment data
```{r}
unemployment_df=
  read_csv("./data/fivethirtyeight_datasets/unemployment.csv")%>%
  janitor::clean_names()%>%
  pivot_longer(jan:dec, names_to = "month", values_to = "unemployment") %>%
  mutate(month = month.abb[as.factor(month)])

```
Merge the datasets
```{r}
pols_snp = 
  left_join(pols_df, snp_df, by=c("year", "month"))
complete_df =
  left_join(pols_snp,unemployment_df, by=c("year", "month"))
```
The original datasets from the FiveThirtyEight data folders contain information about national politicians and their associated parties (democrat or republican) between 1947 and 2015, the closing values in the S&P stock market index between 1950 and 2015, and the unemployment rates by month between 1948 and 2015. The merged dataset contains information on the party the president was given a certain year and month, the number of representatives in each party in the senate and house, as well as the closing values in the S&P stock market index and unemployment percentage by month and year. 

Specific values for each dataset:

Pols-month: `r nrow(pols_df)` rows, `r ncol(pols_df)` variables, between `r min(pols_df$year)` and `r max(pols_df$year)`.

SNP: `r nrow(snp_df)` rows, `r ncol(snp_df)` variables, between `r min(snp_df$year)` and `r max(snp_df$year)`.

Unemployment: `r nrow(unemployment_df)` rows, `r ncol(unemployment_df)` variables, between `r min(unemployment_df$year)` and `r max(unemployment_df$year)`

Complete data: `r nrow(complete_df)` rows, `r ncol(complete_df)` variables, between `r min(complete_df$year)` and `r max(complete_df$year)`.





